---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# boostrq

Package for boosting regression quantiles (Bauer, Haupt, & Linner, 2021). The functionality as well as the implementation was heavily inspired by the great package for model-based boosting [mboost](https://github.com/boost-R/mboost).
Until now, only the linear baselearner (brq) is implemented, the nonlinear baselearner (brqss) will follow soon.

## Installation

You can install the development version of boostrq from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("stefanlinner/boostrq")
```

## Example

The following demonstrates the basic functionality of the package with a simple example.

```{r example}
### Attaching the package to the search path
library(boostrq)

### Fitting your first boosting regression quantiles model.
boosted.rq <- 
  boostrq(
    formula = mpg ~ brq(cyl * hp) + brq(am + wt), # all common formula operators (*,:,^, etc.) can be used in the function brq()
    data = mtcars, 
    mstop = 200, 
    nu = 0.1, 
    tau = 0.5
  )

### Get some basic information about your model 
boosted.rq$formula # the model formula

boosted.rq$nu # the learning rate

boosted.rq$offset # the initialization of the algorithm (default: median of response)

boosted.rq$baselearner.names # names of the baselearners

boosted.rq$call # the model call

boosted.rq$mstop() # current number of iterations

# or use
mstop(boosted.rq)

### Print your fitted model to get a collection of that basic information
boosted.rq

### Selection frequency of each component
boosted.rq$selection.freqs()

### Estimated coefficients for current number of iterations
boosted.rq$coef(aggregate = "sum") # also try aggregate = "cumsum" or "none"

# or use
coef(boosted.rq, aggregate = "sum")

### Printing result summaries 
summary(boosted.rq)

```

```{r, results='hide'}
### Have a look at the underlying baselearner model matrices
boosted.rq$baselearner.matrix()

### Selected component for each iteration
boosted.rq$xselect()
c("Intercept", boosted.rq$baselearner.names)[boosted.rq$xselect() + 1]

### Current working residuals (negative gradients)
boosted.rq$neg.gradients()

### Course of empirical risk during the boosting process
boosted.rq$risk()

### Current fitted values
boosted.rq$fitted()
# or use
fitted(boosted.rq)

### Current residuals
boosted.rq$resid()
# or use
residuals(boosted.rq)

### Model predictions
dat.pred <- data.frame(
  cyl = c(4, 6),
  hp = c(90, 134),
  wt = c(3.125, 2.485), 
  am = c(1, 0)
)
boosted.rq$predict(newdata = dat.pred, aggregate = "sum") # also try aggregate = "cumsum" or "none"
# or use
predict(boosted.rq, newdata = dat.pred, aggregate = "sum")

```

```{r}

### Update the number of iterations without to fully refit the model
### If mstop_new > mstop_current: The fitting process will start at the current number of iterations
### If mstop_new < mstop_current: The model result are subsetted, thus, the model is not refitted

### current number of iterations
boosted.rq$mstop()

### Update number of iterations
boosted.rq <- boosted.rq[300]
boosted.rq$mstop()

# or use
boosted.rq$subset(400)
boosted.rq$mstop()

# or use
mstop(boosted.rq) <- 100
boosted.rq$mstop()

```

